---
title: "MovieMetaData_CaseStudy"
author: "Sopan"
date: "2025-06-06"
output:
  pdf_document: default
  html_document: default
---
```{r}
# Load required packages
library(tidyverse)
library(jsonlite)

# Read the CSV file
movies_metadata <- read_csv("MovieData/movies_metadata.csv")

# Check the structure of your data
glimpse(movies_metadata)

# Clean and expand the genres column
movies_expanded <- movies_metadata %>%
  # Convert JSON string to proper format (replace single quotes with double quotes)
  mutate(genres_clean = str_replace_all(genres, "'", '"')) %>%
  # Parse JSON strings into list columns
  mutate(genres_list = map(genres_clean, ~ fromJSON(.x)$name)) %>%
  # Unnest the list column to create multiple rows
  unnest(genres_list, keep_empty = TRUE) %>%
  # Rename the new column
  rename(genre_name = genres_list) %>%
  # Remove intermediate columns if needed
  select(-genres_clean)

# View the result
head(movies_expanded)

# Save the expanded data if needed
write_csv(movies_expanded, "MovieData/movies_metadata_expanded.csv")
```
# Clean the Data and prepare for Analysis
```{r}
library(tidyverse)
library(lubridate) # For timestamp conversion if needed
```

## Step 1: Load and process ratings data
```{r}
ratings <- read_csv("MovieData/ratings_small.csv") %>%
  # Calculate average rating per movie
  group_by(movieId) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    rating_count = n()
  ) %>%
  ungroup()
```
## Step 2: Load and process movie metadata

```{r}
movies <- read_csv("MovieData/movies_metadata_expanded.csv") %>%
  # Convert id to numeric to match with ratings (some cleaning might be needed)
  mutate(
    movieId = as.numeric(id)  # Assuming 'id' in metadata matches 'movieId' in ratings
  ) %>%
  # Handle any rows that didn't convert properly
  filter(!is.na(movieId))
```


## Step 3: Merge the datasets
```{r}
movies_with_ratings <- movies %>%
  left_join(ratings, by = "movieId") %>%
  # You might want to arrange by rating or other criteria
  arrange(desc(avg_rating))
```



## Step 4: Optional - Add rating categories

```{r}
movies_with_ratings <- movies_with_ratings %>%
  mutate(
    rating_category = case_when(
      avg_rating >= 4.5 ~ "Excellent",
      avg_rating >= 4.0 ~ "Good",
      avg_rating >= 3.0 ~ "Average",
      avg_rating >= 2.0 ~ "Below Average",
      avg_rating >= 0 ~ "Poor",
      TRUE ~ "No Ratings"
    )
  )
```

## Step 5: View and save results

```{r}
glimpse(movies_with_ratings)
```

# Save the merged data
```{r}
write_csv(movies_with_ratings, "MovieData/movies_with_ratings.csv")
```


# Top rated movies
```{r}
top_movies <- movies_with_ratings %>%
  filter(rating_count > 50) %>%  # Only movies with sufficient ratings
  arrange(desc(avg_rating)) %>%
  select(title, avg_rating, rating_count, genres)

head(top_movies, 10)
```

# Visualizations: Movie Recommendation 

## Top-Rated Movies (Weighted by Number of Ratings)

```{r}
movies_with_ratings <- movies_with_ratings %>%
  mutate(weighted_rating = (avg_rating * rating_count) / (rating_count + 100))  # Bayesian average

top_movies <- movies_with_ratings %>%
  filter(rating_count > 30) %>%  # Minimum 30 ratings
  arrange(desc(weighted_rating)) %>%
  head(20)

ggplot(top_movies, aes(x = reorder(title, weighted_rating), y = weighted_rating, fill = weighted_rating)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top-Rated Movies (Adjusted for Rating Count)", 
       x = "Movie Title", 
       y = "Weighted Average Rating") +
  scale_fill_gradient(low = "#ff9999", high = "#66b3ff") +
  theme_minimal()
```
## Most Popular Genres by Average Rating
```{r}
genre_ratings <- movies_with_ratings %>%
  separate_rows(genres, sep = "\\}, \\{") %>%  # Split genres
  mutate(genre = str_extract(genres, "'name': '[A-Za-z]+")) %>%
  mutate(genre = gsub("'name': '", "", genre)) %>%
  filter(!is.na(genre)) %>%
  group_by(genre) %>%
  summarise(
    avg_rating = mean(avg_rating, na.rm = TRUE),
    movie_count = n()
  ) %>%
  filter(movie_count > 50)  # Only genres with sufficient data

ggplot(genre_ratings, aes(x = reorder(genre, avg_rating), y = avg_rating, fill = avg_rating)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Highest-Rated Movie Genres", 
       x = "Genre", 
       y = "Average Rating") +
  scale_fill_gradient(low = "#ffcc99", high = "#ff6666") +
  theme_minimal()
```
## Rating Distribution by Genre
```{r}
movies_with_genres <- movies_with_ratings %>%
  separate_rows(genres, sep = "\\}, \\{") %>%
  mutate(genre = str_extract(genres, "'name': '[A-Za-z]+")) %>%
  mutate(genre = gsub("'name': '", "", genre)) %>%
  filter(!is.na(genre) & !is.na(avg_rating))

ggplot(movies_with_genres, aes(x = genre, y = avg_rating, fill = genre)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Rating Distribution by Genre", 
       x = "Genre", 
       y = "Average Rating") +
  theme_minimal() +
  theme(legend.position = "none")
```
## Movie Recommendations Based on User Preferences
```{r}
# Example: User likes "The Dark Knight" (assuming movieId = 155)
target_movie <- "The Dark Knight"

# Find similar movies by genre and rating
similar_movies <- movies_with_ratings %>%
  filter(str_detect(genres, "Action") &  # Same genre
         avg_rating >= 4.0 &             # Highly rated
         title != target_movie) %>%      # Exclude the target
  arrange(desc(avg_rating)) %>%
  head(10)

ggplot(similar_movies, aes(x = reorder(title, avg_rating), y = avg_rating, fill = avg_rating)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = paste("Movies Similar to", target_movie), 
       x = "Movie Title", 
       y = "Average Rating") +
  scale_fill_gradient(low = "#99cc99", high = "#006600") +
  theme_minimal()
```


